Nitric-oxide-feedback to ciliary photoreceptor cells orchestrates UV avoidance

### 2023-01-24 text/Manuscript.Rmd
edit Methods

cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018]. In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS). The fluorescence intensity of DAF-FM, which specifically detected INNOS, showed a transient tonic-like increase dependent on UV stimulation of the cPRCs. The graph shows a decrease in intensity after UV light is switched off, but the response of DAF-FM to NO is irreversible and this decrease is due to breaching by the relatively strong excitation light. Attempts to measure changes in the intensity of the calcium indicator RGECO1a simultaneously with DAF-FM imaging did not lead to detection of changes in fluorescence intensity compared to GCaMP6s and could only visualise the cPRCs. In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs. This tonic response may involve dopaminergic nerves based on single-cell data (Fig) [@puopolo2001][@zhang2007][@Randel2014a]. In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018]. These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.

@@

To identify GPR173 receptors, we obtained a database of vertebrate SREB sequences, including GPR173, GPR85 and GPR27 from (Breton et al. 2021). From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 49 species investigated. HMM models were run in HMMR3 with an e-value of 1e−15. The same SREB sequences were used to carry out similarity-based searches using BlastP with an e-value cutoff of 1e−15. We merged these two databases and ran CD-Hit (Fu et al. 2012) to eliminate redundant sequences (at a 99% threshold). To identify the sequences that are closely related to the GPR173 sequences, we ran a cluster-based analysis in CLANS (Frickey and Lupas 2004).The CLANS analysis is available as supplementary fig. S4, Supplementary Material online. To identify clusters, we used the convex-clustering option with 100 jack-knife replicates. The SREB receptors are extremely well conserved and form an easily recognizable cluster. To analyze the phylogeny of SREB receptors, the cluster containing these receptors together with monoaminergic receptors were parsed and used for tree building. We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al. 2009). To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al. 2020). The sequences used for the phylogenetic analysis are available in supplementary XX, Supplementary Material online, the trimmed alignment is available in supplementary XX, Supplementary Material online. The raw nexus tree of the SREB receptors is available in supplementary file XX, Supplementary Material online.

>>>>

cPRCs expressing UV-absorbing c-opsin1 have high resting calcium levels and show hyperpolarising and subsequent depolarising responses to UV light (405 nm) [@Veraszto2018]. In the present study, we have shown that the cells that strongly synapse with cPRCs are interneurons that specifically express NOS (INNOS). The fluorescence intensity of DAF-FM, which specifically detected INNOS, showed a transient tonic-like increase dependent on UV stimulation of the cPRCs. The graph shows a decrease in intensity after UV light is switched off, but the response of DAF-FM to NO is irreversible and this decrease is due to breaching by the relatively strong excitation light. Attempts to measure changes in the intensity of the calcium indicator RGECO1a simultaneously with DAF-FM imaging did not lead to detection of changes in fluorescence intensity compared to GCaMP6s and could only visualise the cPRCs. In calcium imaging using GCaMP6s, INNOS identified by the positional relationship between the cPRC and INNOS and the labelling of RYa depolarised in clear synchrony with the hyperpolarising response of the cPRCs. This tonic response may involve dopaminergic nerves based on single-cell data (Figure 6A) [@puopolo2001][@zhang2007][@Randel2014a]. In the mammalian retina, it is known that signals from bipolar photoreceptor cells are transmitted to downstream NOS-expressing amacrine cells (NOAC), where NO is produced [@jacoby2018]. These suggest that UV stimulation of cPRCs causes the production of NO via a depolarising response of INNOS.

@@

To identify NIT-GCs, we obtained a "seed" database of Adenylate and Guanylate cyclase catalytic domain in Pfam database, PF00211. From these sequences, we produced a Hidden Markov Model (HMM) and used this to mine the 45 metazoan species, 2 choanoflagellate species and 2 filasterea species investigated. HMM models were run in HMMR3 with an e-value of 1e−15. We ran CD-Hit (Fu et al. 2012) to eliminate redundant sequences (at a 80% threshold). The CLANS analysis is available as **Supplementary File 1**. To identify clusters, we used the convex-clustering option with 100 jack-knife replicates. The NIT-GCs are extremely well conserved in membrane-bound guanylate cyclases and form an easily recognizable cluster. To analyze the phylogeny of NIT-GCs, the cluster containing these GCs together with membrane-bound guanylate cyclases were parsed and used for tree building. We aligned the sequences with MAFFT version 7, with the iterative refinement method E-INS-i. Alignments were trimmed with TrimAl in gappy-out mode (Capella-Gutiérrez et al. 2009). To calculate maximum-likelihood trees, we used IQ-tree2 with the LG+G4 model. To calculate branch support, we ran 1,000 replicates with the aLRT-SH-like and aBayes methods (Minh et al. 2020). The sequences used for the phylogenetic analysis are available in **Supplementary File 2**, the trimmed alignment is available in **Supplementary File 3**.
